---
title: "Data story project"
filters:
  - shinylive
---

### What contributes to the difference in wait times for admission to wards in local public hospitals?

The CNA had reported in April 2023 that the median waiting time for ward admissions had increased from about 5 hours to 7.2 hours during that period^[1](https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136)^. The health ministry had explained that it was due to the rise in number of older patients with more complex medical needs requiring longer hospitalisation^[1](https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136)^. Due to the ageing population in Singapore, long waiting time for admission to wards has been a long-standing issue over the years, and it was further exacerbated by the Covid-19 pandemic^[2](https://www.channelnewsasia.com/singapore/hospitals-beds-waiting-time-50-hours-admission-ng-teng-fong-sengkang-3014596),[3](https://www.straitstimes.com/singapore/health/most-public-hospitals-saw-more-ae-patients-in-past-week-compared-with-past-month)^.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyr)
library(readr)
library(leaflet)
```

``` {=html}
<iframe height="1100" width="100%" frameborder="no" src="https://bocongg.shinyapps.io/proj_app/"> </iframe>
```


#### Bar chart
```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
final_dataset %>% 
  filter(Month == "01",
         Day == "01") %>%
    ggplot(aes(x = occ_rate_num)) + 
    geom_col(aes(y = fct_reorder(hospital, desc(hospital)),
                 fill = hospital)) +
    labs(title = "Bed Occupancy Rate (BOR)",
         subtitle = "01/01/2023",
         x = "BOR (%)",
         y = "Hospital",
         caption = "Source: Ministry of Health") +
    geom_text(aes(occ_rate_num, hospital, label = occ_rate_num),
              hjust = 1.3) +
    theme(legend.position = "none")

final_dataset %>% 
  filter(Month == "01",
         Day == "01") %>%
    ggplot(aes(x = attendance)) + 
    geom_col(aes(y = fct_reorder(hospital, desc(hospital)),
                 fill = hospital)) +
    labs(title = "Attendance at Emergency Medicine Department",
         subtitle = "01/01/2023",
         x = "No. of patients",
         y = "Hospital",
         caption = "Source: Ministry of Health") +
    geom_text(aes(attendance, hospital, label = attendance),
              hjust = 1.3) +
    theme(legend.position = "none")
```

#### Calendar Heatmap
```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
calendar_plot <- final_dataset %>%  
  filter(hospital == "AH") %>%
    ggplot(aes(x = fct_inorder(Day_wk),
               y = monthweek,
               fill = wait_duration,
               text = paste("Waiting time:", wait_duration, "hrs"))) + 
    geom_tile(colour = "white") +
    geom_text(aes(label = Day),
              color = "black",
              size = 2) +
    facet_wrap( ~ fct_inorder(fullmonth),
                ncol = 3,
                scales = "fixed") +
    scale_fill_gradient(low="green", high = "red") +
    labs(x = "Alexandra Hospital (2023)",
         y = "",
         title = "Waiting Time for Admission to Ward",
         subtitle = "Alexandra Hospital (2023)",
         fill = "Hours (h)",
         caption = "Source: Ministry of Health") +
    theme(axis.text.y=element_blank(), 
          axis.ticks.y=element_blank()) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme(panel.spacing.y = unit(1.1, "cm"))

ggplotly(calendar_plot, tooltip = "text") %>%
  config(displayModeBar = FALSE) %>%
  layout(dragmode = FALSE)
```

#### Line + bar chart
```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
final_dataset %>% 
  filter(Month == "01",
         hospital == "AH") %>%
    ggplot(aes(x = Dates)) + 
    geom_col(aes(y = occ_rate_num/10), fill = "lightgrey") +
    geom_point(aes(y = wait_duration), size = 2) + 
    geom_line(aes(y = wait_duration, group = hospital, color = hospital), size = 0.75) + 
    labs(title = "Waiting Time for Admission to Ward & Bed Occupancy Rate (BOR)",
         subtitle = "Alexandra Hospital (Jan, 2023)",
         x = "Date",
         y = "Hours (h)",
         caption = "Source: Ministry of Health") +
    theme(legend.position = "none") +
    scale_y_continuous(sec.axis = sec_axis(~.*10, name = "BOR (%)")) +
    scale_x_discrete(guide = guide_axis(angle = 45))

final_dataset %>% 
  filter(Month == "01",
         hospital == "AH") %>%
    ggplot(aes(x = Dates)) + 
    geom_col(aes(y = attendance/10), fill = "lightgrey") +
    geom_point(aes(y = wait_duration), size = 2) + 
    geom_line(aes(y = wait_duration, group = hospital, color = hospital), size = 0.75) + 
    labs(title = "Waiting Time for Admission to Ward & EMD Attendance",
         subtitle = "Alexandra Hospital (Jan, 2023)",
         x = "Date",
         y = "Hours (h)",
         caption = "Source: Ministry of Health") +
    theme(legend.position = "none") +
    scale_y_continuous(sec.axis = sec_axis(~.*10, name = "EMD Attendance")) +
    scale_x_discrete(guide = guide_axis(angle = 45))
```
#### Scatterplot
```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
final_dataset %>%
  filter(Month == "01",
         hospital == "AH") %>%
  ggplot(aes(wait_duration, occ_rate_num)) +
  geom_point() +
  geom_smooth(method = "lm", se=F) +
  labs(title = "Waiting Time for Admission to Ward & Bed Occupancy Rate (BOR)",
         subtitle = "Alexandra Hospital (Jan, 2023)",
         x = "Hours (h)",
         y = "BOR (%)",
         caption = "Source: Ministry of Health")

final_dataset %>%
  filter(Month == "01",
         hospital == "AH") %>%
  ggplot(aes(wait_duration, attendance)) +
  geom_point() +
  geom_smooth(method = "lm", se=F) +
  labs(title = "Waiting Time for Admission to Ward & EMD attendance",
         subtitle = "Alexandra Hospital (Jan, 2023)",
         x = "Hours (h)",
         y = "No. of patients",
         caption = "Source: Ministry of Health")
```

```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
final_dataset %>%
  filter(Month == "01",
         Dates == "01/01/2023") %>%
    ggplot(aes(x = "", y = attendance, fill = factor(hospital))) + 
    geom_bar(width = 1, stat = "identity", color = "white") + 
    theme(axis.line = element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    labs(title = "Attendances at Emergency Medicine Department, 1st January (2023)",
         x = NULL,
         y = NULL,
         fill = "Hospital",
         caption = "Source: Ministry of Health") +
    coord_polar(theta = "y", start = 0) +
    geom_text(aes(label = attendance),
              position = position_stack(vjust = 0.5),
              size = 5) +
    theme_void()


```


```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
df <- read.csv(textConnection(
  "Name,Lat,Long
  Singapore General Hospital (SGH),103.8348,1.2804
  Changi General Hospital(CGH),103.9496,1.3402
  Sengkang General Hospital (SKH),103.8931,1.3943
  National University Hospital (NUH),103.7831,1.2939
  Ng Teng Fong General Hospital (NTFGH),103.7439,1.3350
  Alexandra Hospital (AH),103.8013,1.2866
  Tan Tock Seng Hospital (TTSH),103.8458,1.3214
  Khoo Teck Puat Hospital (KTPH),103.8394,1.4241"))

leaflet(df) %>%
  addTiles() %>%
   addMarkers(~Long, ~Lat, label = ~htmlEscape(Name)) %>%
    setView(lng = 103.8, lat = 1.366667, zoom = 12)
```